---
title: "DataCamp"
author: "Robert Stroud"
date: "10/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", highlight = TRUE)
```

# Bootstrapping for estimating for a parameter

##Generate bootstrap distribution for median

When building a bootstrap distribution for a single statistic, we first generate a series of bootstrap resamples, and then record the relevant statistic (in this case, the median) of each distribution.

In this exercise, the tidyverse packages, including ggplot2 have been loaded for you, as has infer.

INSTRUCTIONS
100 XP
Fill in the blanks in the code to generate 15,000 bootstrap distributions of rent in the manhattan data frame and record the median of each bootstrap distribution.
Save the resulting data frame of 15,000 bootstrap medians as rent_ci_med.
View the str()-ucture of rent_ci_med to get a sense of what this object looks like.
Visualize the bootstrap distribution of medians with a histogram of binwidth = 50, and describe the shape, center, and spread of the bootstrap distribution of medians.

```{r}
library(tidyverse, ggplot2, infer)
manhattan <- read.csv("https://assets.datacamp.com/production/course_5103/datasets/manhattan.csv")

# Generate bootstrap distribution of medians
rent_ci_med <- manhattan %>%
  # Specify the variable of interest
  specify(response = rent) %>%  
  # Generate 15000 bootstrap samples
  generate(reps = 15000, type = "bootstrap") %>% 
  # Calculate the median of each bootstrap sample
  calculate(stat = "median")

# View the structure of rent_ci_med
str(rent_ci_med)

# Plot a histogram of rent_ci_med
ggplot(manhattan, aes(x = rent_ci_med)) +
  geom_histogram()
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
