---
title: "ClassNotes"
author: "Robert Stroud"
date: "10/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Supervised Learning:
y - repsonse, dependent 
x - predictors, independent, inputs

use a scatterplot to compare two numeric variables. 

Describe three things

Form - linear / non linear ex. 
Direction - positive / negative
Strength - Strong, moderate, weak ex. most subjective of the three. 

When creating function to model some data, explanation and prediction are potential focuses. 
Prediction is significantly more useful than explanation. 

an estimation funciton will have a carat/hat above it. 
ex. y = f(x) + Error, f(x) is the signal, error is the noise. 

We want to minimize the noise. 

Ei = y - f(x)

wnat to minimize the sum of squares error. 

f(x) = y = b_0 + b_1(X_i)
b_0 - y intercept
b_1 - slope 

sum of squares = Sum of (y_i - b_0 - b_iX_i)^2

take the derivative of above. 

set it to zero
solve for whatever term was used for derivative. 

it ends up being in repsect to b_0
b_0 = avg(y) - b_1*avg(x)
b_1 = sum of (x_1 - avg(x))(y_i - avg(y)) / sum of (x_i - avg(x))^2

these two equations result in the minimization of the sum of square eirs. 

r - correlation coefficient 
 -1 < r < 1
 perfect negative relationship will be a -1. 
 perfect postive relationship will be 1. 

```{r}
library(MASS)
library(dplyr)
library(ggplot2)
?Boston
##geom smooth used to give a linear model, lm stands for linear model. 
## cor() is the method to find the correlation coefficient. 
##lm(response ~ predictors, data = DF), this describe data, give intercpet and slope. 
## sumamry of mod gives several values incuding r squared. taking the sqrt of that will give the correlation coeffiecient. 

ggplot(Boston, aes(x = rm, y = medv)) + geom_point() + geom_smooth(method = "lm")
cor(Boston$medv, Boston$rm)
mod <- lm(medv ~ rm, data = Boston)
mod
summary(mod)
summary(mod)$r.squared
```

imbedding this into an rmarkdown document is valuable. because
```{r}
library(ISLR)
##regress balance, putting a one after the . will do regress over intercept. 
mod2 <- lm(Balance~., data = Credit)
```
Putting the code in the document is preferred, so if the data changes the document will change dynamically. 

The $R^2_{adj}=`r summary(mod2)$adj`$ 

Backwards Elimination - regressing the response on to evey predictor. Then get rid of the variable that is the least significant, p-value / f-value. This process continues until all variables in the model are significant. 

Forward Selection - start with one variable, create several regression models, then select the most significant. Then select another variable and create several 2-variable regression models and select the most significant, this continues. 

```{r}
##FORWARD selection model example. 
##SCOPE selects all the variable that can be used in the test, add1 is the function used to test. 
## looking at the f-value because of test = "F", and we are looking for the largest f value. 
## choose the variable based on largest f-value and p-value.
library(ISLR)
mod.fs <- lm(Balance ~ 1, data = Credit)
SCOPE <- (~Income + Limit + Rating + Cards + Age + Education + Gender + Student + Married + Ethnicity)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Rating)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Income)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Student)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Limit)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Cards)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Age)
add1(mod.fs, scope = SCOPE, test = "F")
summary(mod.fs)
```

Backward Elimination
```{r}
## using the dot says over the whole data frame, instead of just the intercept.
##drop 1 used to test. 
mod.be <- lm(Balance ~ ., data = Credit)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - Education)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - Ethnicity)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - Married)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - ID)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - Gender)
drop1(mod.be, test = "F")
summary(mod.be)
```

All-in-One Function
```{r}
## the direction gives which type, forward means forward selction. 
stepAIC(lm(Balance ~ 1, data = Credit), scope = SCOPE, 
        direction = "forward", test = "F")

null <- lm(Balance ~1, data = Credit)
full <- lm(Balance ~ ., data = Credit)
##forward selection
stepAIC(null, scope = list(lower = null, upper = full),
        direction = "forward", test = "F")
##backward elimination
stepAIC(full, scope = list(lower = null, upper = full),
        direction = "backward", test = "F")
```
Standard Normal Score
z Score = any statistic minus the mean of that statistic over the standard deviation. 
t score 

if you have a random variable with a distribution, provided n ~ sample size, is sufficiently large ~ depends on the sample, Xbar is the sample mean. distribution of Xbar has an approximate normal shape. If you have a uniform distribution, if you take a sufficiently large sample from it, and take the mean it will provide a bell curve.

for the graph of z, there shouuld be alpa / 2 area in the tails, the area to the left is called z_alpha / 2 and the one to the right is z_1-alpha / 2.

p is the cumalalitve districbution function. 
q will give us the invverse cumalitive distribution funtion, 

Confidence Interval of mew of probablity statment of pt estimate minus the margin of error, pt estimate plus the margin of error. 
once computed the confience intervale talks about the percentage of the true mean is in the calculated range. 
```{r}
junk <- runif(10000, 0, 10)
hist(junk)
sims <- 10000
xbar <- numeric(sims)
for(i in 1:sims){
  x <- runif(16, 0, 10)
  xbar[i] <- mean(x)
}
hist(xbar)
mean(xbar)
sd(xbar)
qnorm(.95)
```