---
title: "ClassNotes"
author: "Robert Stroud"
date: "10/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Supervised Learning:
y - repsonse, dependent 
x - predictors, independent, inputs

use a scatterplot to compare two numeric variables. 

Describe three things

Form - linear / non linear ex. 
Direction - positive / negative
Strength - Strong, moderate, weak ex. most subjective of the three. 

When creating function to model some data, explanation and prediction are potential focuses. 
Prediction is significantly more useful than explanation. 

an estimation funciton will have a carat/hat above it. 
ex. y = f(x) + Error, f(x) is the signal, error is the noise. 

We want to minimize the noise. 

Ei = y - f(x)

wnat to minimize the sum of squares error. 

f(x) = y = b_0 + b_1(X_i)
b_0 - y intercept
b_1 - slope 

sum of squares = Sum of (y_i - b_0 - b_iX_i)^2

take the derivative of above. 

set it to zero
solve for whatever term was used for derivative. 

it ends up being in repsect to b_0
b_0 = avg(y) - b_1*avg(x)
b_1 = sum of (x_1 - avg(x))(y_i - avg(y)) / sum of (x_i - avg(x))^2

these two equations result in the minimization of the sum of square eirs. 

r - correlation coefficient 
 -1 < r < 1
 perfect negative relationship will be a -1. 
 perfect postive relationship will be 1. 

```{r}
library(MASS)
library(dplyr)
library(ggplot2)
?Boston
##geom smooth used to give a linear model, lm stands for linear model. 
## cor() is the method to find the correlation coefficient. 
##lm(response ~ predictors, data = DF), this describe data, give intercpet and slope. 
## sumamry of mod gives several values incuding r squared. taking the sqrt of that will give the correlation coeffiecient. 

ggplot(Boston, aes(x = rm, y = medv)) + geom_point() + geom_smooth(method = "lm")
cor(Boston$medv, Boston$rm)
mod <- lm(medv ~ rm, data = Boston)
mod
summary(mod)
summary(mod)$r.squared
```

imbedding this into an rmarkdown document is valuable. because
```{r}
library(ISLR)
##regress balance, putting a one after the . will do regress over intercept. 
mod2 <- lm(Balance~., data = Credit)
```
Putting the code in the document is preferred, so if the data changes the document will change dynamically. 

The $R^2_{adj}=`r summary(mod2)$adj`$ 

Backwards Elimination - regressing the response on to evey predictor. Then get rid of the variable that is the least significant, p-value / f-value. This process continues until all variables in the model are significant. 

Forward Selection - start with one variable, create several regression models, then select the most significant. Then select another variable and create several 2-variable regression models and select the most significant, this continues. 

```{r}
##FORWARD selection model example. 
##SCOPE selects all the variable that can be used in the test, add1 is the function used to test. 
## looking at the f-value because of test = "F", and we are looking for the largest f value. 
## choose the variable based on largest f-value and p-value.
library(ISLR)
mod.fs <- lm(Balance ~ 1, data = Credit)
SCOPE <- (~Income + Limit + Rating + Cards + Age + Education + Gender + Student + Married + Ethnicity)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Rating)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Income)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Student)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Limit)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Cards)
add1(mod.fs, scope = SCOPE, test = "F")
mod.fs <- update(mod.fs, .~. + Age)
add1(mod.fs, scope = SCOPE, test = "F")
summary(mod.fs)
```

Backward Elimination
```{r}
## using the dot says over the whole data frame, instead of just the intercept.
##drop 1 used to test. 
mod.be <- lm(Balance ~ ., data = Credit)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - Education)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - Ethnicity)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - Married)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - ID)
drop1(mod.be, test = "F")
mod.be <- update(mod.be, .~. - Gender)
drop1(mod.be, test = "F")
summary(mod.be)
```

All-in-One Function
```{r}
## the direction gives which type, forward means forward selction. 
stepAIC(lm(Balance ~ 1, data = Credit), scope = SCOPE, 
        direction = "forward", test = "F")

null <- lm(Balance ~1, data = Credit)
full <- lm(Balance ~ ., data = Credit)
##forward selection
stepAIC(null, scope = list(lower = null, upper = full),
        direction = "forward", test = "F")
##backward elimination
stepAIC(full, scope = list(lower = null, upper = full),
        direction = "backward", test = "F")
```
Standard Normal Score
z Score = any statistic minus the mean of that statistic over the standard deviation. 
t score 

if you have a random variable with a distribution, provided n ~ sample size, is sufficiently large ~ depends on the sample, Xbar is the sample mean. distribution of Xbar has an approximate normal shape. If you have a uniform distribution, if you take a sufficiently large sample from it, and take the mean it will provide a bell curve.

for the graph of z, there shouuld be alpa / 2 area in the tails, the area to the left is called z_alpha / 2 and the one to the right is z_1-alpha / 2.

p is the cumalalitve districbution function. 
q will give us the invverse cumalitive distribution funtion, 

Confidence Interval of mew of probablity statment of pt estimate minus the margin of error, pt estimate plus the margin of error. 
once computed the confience intervale talks about the percentage of the true mean is in the calculated range. 
```{r}
junk <- runif(10000, 0, 10)
hist(junk)
sims <- 10000
xbar <- numeric(sims)
for(i in 1:sims){
  x <- runif(16, 0, 10)
  xbar[i] <- mean(x)
}
hist(xbar)
mean(xbar)
sd(xbar)
qnorm(.95)
```

WEEK 9
In Class Problems:

Bayes #1:

Dans's Diner employes three dishwashers. Al washes 20% of the dishes and breaks only 1% of those he handles. Betty and Chuck each wash 40% of the dishes and betty breaks only 1% of hers and chuck breaks 5% of the dhises he washes. You go to Dans for supper one night and hear a dish break in the sink. Whats the probability that chuck is on the job.

A = event Al is on the job
p(A) = 0.20
p(BD(A)) = 0.01
p(~BD(A)) = 0.99
A = 0.002
~A = 0.198

B = event Betty is on the job
p(B) = 0.40
p(BD(B)) = 0.01
p(~BD(B)) = 0.99
B = 0.004
~B = 0.396

C = event Chuck is on the job
p(C) = 0.40
p(BD(C)) = 0.05
p(~BD(C)) - 0.95
C = 0.02
~C = 0.38

p(BD(C)) = p(c&BD) / P(BD) = p(BD(C)) * P(c) / P(BD(C)) * P(C) + p(BD(B)) * P(B) + P(BD(A)) * P(A)
P(BD(C)) = 0.2 / 0.02 + 0.004 + 0.002 = 0.769

Bayes 2:
Nicole is fliying from boston to denver with a connection in chigace. the probability that her flight levaves on time is 0.25. if the fligh is on time the probability that her luggage will make the connection flightis 0.85 but if the flight is delayed the probability that the luggage will make it is only 0.55

Suppose you pick her up at the denver airpoirt with her luggage is ont there. What is the probabality that nicoles first flight was delayed. Ans = 0.9

P(OT) = 0.25
P(LA(OT)) = 0.85
p(~LA(OT)) = 0.15

P(~OT) = 0.75
P(LA(~OT)) = 0.55
p(~LA(~OT)) = 0.45

P(~OT(~LA)) = P(~LA(~OT)) * P(~OT) / P(~LA(~OT)) * P(~OT) + P(~LA(OT)) * P(OT)
P(~OT(~LA)) = (0.45)(0.75) / (0.45)(0.75) + (0.15)(0.25)
P(~OT(~LA)) = 0.9

Expected Values 1:
You draw a card from a deck. If you get a red card, you win nothing. If you get a spade, you win $8. For any club, you get $20 plus an extra $40 for the ace of clubs.

Create a probability model for the amount you win at this game.
Find the expected amount you'll win. ($7.77)
How much would you be willing to pay to play this game?

P(0) = 26 / 52
P(8) = 13 / 52
p(20) = 12 / 52
p(60) = 1 / 52

e(x) = sum of(x * P(x =x)) = 0 * (26/52) + 8 * (13 / 52) + 20 (12 / 52) + 60 (1 / 52)

for r:
put each x into a seperate vector. 
x for value
px for probability
then create a new vector with sum(X*Px)

variation for discreate variable:
V(x) = theta^2_x = sum of (x - e(x))^2 * P(X=x)

for r(continued):
vx <- sum((x-EX)^2 * px)
SX <- sqrt(VX)

Expected Value 2: 

A couple plans to have children until they get a girl, but they agree they will not have more than three children, even if all are boys. Assume that the probability of having a girl is 53.00%.

Create a probability model for the number of children they'll have.
Find the expected number of children. (1.6909)
Find the expected number of boys they'll have. (0.7947)

p(b) = 0.47
p(g) = 0.53

(Use same rcode as #1 just substitue below values in)
x = # of children
x  P(x)
1  0.53
2  0.47 * 0.53
3  0.47^3 + 0.47^2 * 0.53

y = # of boys
y  p(y)
0 0.53
1 0.47 * 0.53
2 0.47^2 * 0.53
3  0.47^3

r:
y <- 0:3
py <- c(0.53, 0.47*0.53, 0.47^2*0.53, 0.47^3)
EY <- sum(y * py)

Standard Deviation 1: 

You draw a card from a deck. If you get a red card, you win nothing. If you get a spade, you win $13. For any club, you win $26 plus an extra $15 for the ace of clubs. Find the standard deviation of the amount you might win drawing a card.

The standard deviation is_____ ($11.39)


```{r}
x <- c(0, 13, 26, 41)
px <- c(26/52, 13/52, 12/52, 1/52)
EX <- sum(x*px)
VX <- sum((x-EX)^2 * px)
SX <- sqrt(VX)
SX

```

Expected Value and Variance 1:

An insurance policy costs $110 and will pay policyholders $9,000 if they suffer a major injury (resulting in hospitalization) or $2000 if they suffer a minor injury (resulting in lost time from work). The company estimates that 1 in every 2083 policyholders will suffer a major injury and that 1 in 495 will suffer a minor injury.

Create a probability model for the profit on a policy.
What's the company's expected profit on the policy? ($101.64)
What's the standard deviation? ($216.5632)

profit  p(profit)
110     1 - (1/2083 + 1/495)
-1890   1 / 495
-8890   1 / 2083

```{r}
x <- c(110, -1890, -8890)
px <- c(1 - (1/2083 + 1/495), 1/495, 1/2083)
EX <- sum(x*px)
VX <- sum((x - EX)^2 * px)
SX <- sqrt(VX)
EX
SX
```

Coffee Problem:

At a certain coffee shop, all the customers buy a cup of coffee and some also buy a doughnut. The shop owner believes that the number of cups he sells each day is normally distributed with a mean of 320 cups and a standard deviation of 22 cups. He also believes that the number of doughnuts he sells each day is independent of the coffee sales and is normally distributed with a mean of 160 doughnuts and a standard deviation of 16.

The shop is open every day but Sunday. Assuming day-to-day sales are independent, what's the probability he'll sell over 2000 cups of coffee in a week? (0.069)

Let C = # of cups of coffee sold
let D = # of donuts sold
C ~ N(320, 22)
D ~ N(160, 16)

C_Total = C_1 + C_2 + C_3 + C_4 + C_5 + C_6

E[C_Total] = E[C_1 + ... C_6]

E[C_Total] = 6 * 320 = 1920

V[C_Total] = 6 * 22^2 = 

C_Total ~ N(1920, sqrt(6) * 22)

p(C_Total > 2000) = 1 - p(C_Total <= 2000) = 1 - pnorm(2000, 1920, sqrt(6) * 22)

```{r}
P <- 1 - pnorm(2000, 1920, sqrt(6) * 22)
P
```

Coffee Problem pt.2:

if he makes a profit of 50 cents on each cup of coffee and 40 cents on each doughnut, can he reasonably expect to have a day's profit of over $300? Explain. (No. $300 is more than 5 standard deviations above the mean.)

What's the probability that on any given day he'll sell a doughnut to more than half of his coffee customers? (0.5)

profit = 0.5C + 0.4D

C ~ N(320, 22)
D ~ N(160, 16)

E(profit) = 0.50E(c) + 0.40E(d) = 0.50 * 320 + 0.40 * 160 = 224
V(profit) = 0.50^2V(c) + 0.40V(D) = 0.50^2 * 22^2 + 0.40^2 * 16^2 = 
SD(profit) = sqrt(V(profit)) = 12.73

profit ~ N(224, 12.73)

Pregnancies Problem: 

Assume that the duration of human pregnancies can be described by a normal model with mean 265 days and standard deviation 15 days. Answer the following questions.

What percentage of pregnancies should last between 275 and 290 days? (20.5%)

At least how many days should the longest 20% of all pregnancies last? (277.6)

PL ~ N(265, 15)

P(275 <= PL <= 290) = P(P(PL <= 290) - P(PL <= 275))
```{r}
P <- pnorm(290, 265, 15) - pnorm(275, 265, 15)
P
P_1 <- qnorm(.80, 265, 15)
P_1
```

Pregnancy Problem pt2:

Suppose a certain obstetrician is currently providing prenatal care to 90 pregnant women. Let y¯ represent the mean length of their pregnancies. According to the central limit theorem, what is the mean and standard deviation SD(y¯) of the normal model of the distribution of the sample mean, y¯? The mean is (265). SD(y¯) = 1.58.

What is the probability that the mean duration of these patients' pregnancies will be less than 267 days? (0.897)

```{r}
P <- 15 / sqrt(90)
P
P_1 <- pnorm(267, 265, 15)
P_1
```

Create a function that finds the expected value, variance, standard deviation
```{r}
y <- c(110, -1890, -8890)
py <- c(1 - (1/2083 + 1/495), 1/495, 1/2083)

EVS <- function(x, px){
EX <- sum(x*px)
VX <- sum((x - EX)^2 * px)
SX <- sqrt(VX)
print(VX)
print(EX)
SX
}

P <- EVS(y, py)
P
```

Question 2:


```{r}

```